Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.

### 2010-05-02 paper.tex
Fix equilibrium expression for the single fluid free energy model.
Fix wxx, wyy and wxy coefficients in the paper.
Added the Binary liquid test description. Need to calculate profiles. Battery runs out..

\subsection{Shan-Chen model}
In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends
on all local neighbours characteristics:
\begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
\bm{F}(\bm{x})=-G\psi(\rho(\bm{x}))\sum_{i}{w_i \psi(\rho(\bm{x}+\bm{c_i}))\bm{c_{i}}},
\end{equation}
where $\psi$ is the function of local characteristics of a node, and taken
as $\psi(\rho)=1-\exp(-\rho)$. Coefficient $G$ is responsible for the attraction
force between the molecules. Overall, the Shan-Chen force is the approximation of
the force between molecules \cite{kwok,kwok-contact-angle}. The force modifies
the momentum fluxes as:
\begin{equation*}
P_{\alpha\beta}=\biggl(c_s^2\rho+\frac{G}{6}\psi^2+\frac{G}{36}%
|\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
18}\partial_{\alpha}\psi\partial_{\beta}\psi,
\end{equation*}
where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book}
to the van-der-Waals type equation of state:
\begin{equation}
P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
\end{equation}
The lattice Boltzmann simulation for the Shan-Chen model is implemented through the standard implementation (\ref{standard:implementation}). In the presence of the external force the macroscopic velocity is changed as $\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}+\dfrac{F_{\alpha}}{2}$. The force population is defined as \cite{guo}:
\begin{equation}
F_i=w_i\biggl(1-\frac{\omega}{2}\biggr)\biggl(3 (c_{i\alpha}-u_{\alpha}) + 9 (c_{i\beta}u_{\beta})c_{i\alpha}\biggr)F_{\alpha}.
\end{equation}
The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix~\ref{sec:binary_shan_chen}.
\subsection{Free-energy model}
The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation. In comparison with the Shan-Chen model, the free-energy model is simulated through the special choice of the equilibrium distribution function. The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
\begin{equation}
\begin{aligned}
f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
\end{aligned}
\end{equation}
where $w_{ixx}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
$w_{iyy}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
\begin{equation}
\Delta=\frac{1}{6}
\begin{pmatrix}
1&4&1\\
4&-20&4\\
1&4&1\\
\end{pmatrix}\text{ and }
\partial_x=\frac{1}{12}
\begin{pmatrix}
-1&0&1\\
-4&0&4\\
-1&0&1\\
\end{pmatrix}.
\end{equation}

@@

\section{Results}
\label{sec:benchmark}

\section{Conclusion}
%%%% Acknowledgments %%%%%%%%
\section*{Acknowledgments}
A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.

>>>>

\subsection{Shan-Chen model}
In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends
on all local neighbours characteristics:
\begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
\bm{F}(\bm{x})=-G\psi(\rho(\bm{x}))\sum_{i}{w_i \psi(\rho(\bm{x}+\bm{c_i}))\bm{c_{i}}},
\end{equation}
where $\psi$ is the function of local characteristics of a node, and taken
as $\psi(\rho)=1-\exp(-\rho)$. Coefficient $G$ is responsible for the attraction
force between the molecules. Overall, the Shan-Chen force is the approximation of
the force between molecules \cite{kwok,kwok-contact-angle}. The force modifies
the momentum fluxes as:
\begin{equation*}
P_{\alpha\beta}=\biggl(c_s^2\rho+\frac{G}{6}\psi^2+\frac{G}{36}%
|\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
18}\partial_{\alpha}\psi\partial_{\beta}\psi,
\end{equation*}
where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book}
to the van-der-Waals type equation of state:
\begin{equation}
\label{shan:chen:equation:state}
P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
\end{equation}
The lattice Boltzmann simulation for the Shan-Chen model is implemented through the standard implementation (\ref{standard:implementation}). In the presence of the external force the macroscopic velocity is changed as $\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}+\dfrac{F_{\alpha}}{2}$. The force population is defined as \cite{guo}:
\begin{equation}
F_i=w_i\biggl(1-\frac{\omega}{2}\biggr)\biggl(3 (c_{i\alpha}-u_{\alpha}) + 9 (c_{i\beta}u_{\beta})c_{i\alpha}\biggr)F_{\alpha}.
\end{equation}
The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix~\ref{sec:binary_shan_chen}.
\subsection{Free-energy model}
The free-energy or pressure model \cite{swift} allows the incorporation of the
any equation of state to the lattice Boltzmann equation. In comparison with the
Shan-Chen model, the free-energy model is simulated through the special choice
of the equilibrium distribution function. The lattice Boltzmann scheme is
implemented in a standard form (\ref{standard:implementation}), but the
equilibrium function is defined as: \begin{equation}
\begin{aligned}
f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{c_s^2}+\rho \frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr)\\
&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho), \quad i > 0
\end{aligned}
\end{equation}
where $w_{ixx}=\left\{\dfrac{1}{3},-\dfrac{1}{6},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
$w_{iyy}=\left\{-\dfrac{1}{6},\dfrac{1}{3},-\dfrac{1}{6},\dfrac{1}{3}, -\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},-\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
\begin{equation}
\Delta=\frac{1}{6}
\begin{pmatrix}
1&4&1\\
4&-20&4\\
1&4&1\\
\end{pmatrix}\text{ and }
\partial_x=\frac{1}{12}
\begin{pmatrix}
-1&0&1\\
-4&0&4\\
-1&0&1\\
\end{pmatrix}.
\end{equation}

@@

\section{Results}
\label{sec:benchmark}
A few benchmarks were made for comparison purposes between the Shan-Chen and the free-energy model:
\begin{itemize}
\item The first one compares accuracy and performance of calculations of the plane interface between gas and liquid phases.  We included the Shan-Chen equation of state (\ref{shan:chen:equation:state}) to the free-energy model for the purpose of comparison. The domain {\color{red} Michal, please add the domain size} was initialized with the equilibrium densities calculated from Maxwellian area rule for $G=-5.5$. Half domain was initialized with the liquid density $\rho_l=...$, the another half domain was initialized with the gas density $\rho_g=...$. While the prescription of $G$ is enough for the Shan-Chen model to run simulations, the free-energy model operates with the additional parameter $k$ controlling the surface tension. We take values of $k$ {\color{red} Michal, please put a value here.} to guarantee the closeness of calculated profiles to the Shan-Chen model profiles \cite{sbragaglia-contact, kuzmin-thesis}.

While the profiles are not compared since the Shan-Chen model has the density dependent surface tension coefficient \cite{shan-chen-free}, the models are compared in terms of accuracy to reproduce the bulk densities and the computer performance. The following outlines the findings: {\color{red} Michal, please add the benchmarks here}.

\item The binary-liquid models can be compared for the two-phase Poiseuille flow. The flow is driven by pressure gradient implemented as the force, $\frac{\mathrm{d}P}{\mathrm{d}x}$. The walls are implemented as a half-way bounce-back. It is known that BGK fails to reproduce the exact location of the wall. Instead of BGK we used the D2Q9 MRT model \cite{dhumieres-first,dhumieres}. While BGK operates with one parameter $\tau$, the MRT model operates with the number of eigenvalues, i.e. $s_{\rho},s_e,s_{\epsilon},s_{j_x},s_{q_x},s_{j_y},s_{q_y},s_{p_{xx}},s_{p_{xy}}$, giving better stability and proper the Navier-Stokes equation restoration by varying the free model eigenvalues \cite{dhumieres}.  For example for the half-way bounce-back walls representation for the Poiseuille flow the location of the wall is located exactly in the middle between nodes if the energy fluxes eigenvalues are defined as \cite{ginzburg-multireflection}:
\begin{equation}
\label{poiseuille:eigenvalues}
s_{q_{x,y}}=8\frac{2-\omega}{8-\omega},
\end{equation}
where $\omega$ states for the $s_{p_{xx}}$ and $s_{p_{xy}}$ related to the bulk and kinematic viscosities.

The domain {\color{red} Please put size here} is utilized with the two different densities {\color{red} What are they?}. The nodes with $y=0$ and $y=Ny$ represent the half-way bounce-back nodes. The MRT energy-flux eigenvalues are defined through the equation \ref{poiseuille:eigenvalues}, while viscosity related eigenvalues are defined as $s_{p_{xx}}=s_{p_{xy}}=\omega=1$. The velocity profile should obey the following formulae \cite{ginzburg-discontinious}:
\begin{equation}
\text{Need to calculate profiles. Michal, if you already have it - please add.}
\end{equation}

We compare the models in terms of the accuracy for velocity profiles for different density and viscosity ratios and the calculation performance. The following summarizes our findings: {\color{red} Michal, please add here benchmarks}.   
\end{itemize}
\section{Conclusion}
%%%% Acknowledgments %%%%%%%%
\section*{Acknowledgments}
A.~Kuzmin wants to thank the Alberta Ingenuity Fund for their financial support.
